Holomorphic field theories and higher algebra

### 2024-09-26 draft.tex
redid section 2

As a more mundane example, let $\RR\Hol(X,Y)$ denote the derived moduli space of holomorphic maps from some complex manifold $X$ to another complex manifold $Y$.
(For $Y = V$ a vector space, a ``derived'' enhancement is familiar as the higher dimensional cohomology $H^*(X, \cO) \otimes V$.)
In derived geometry one can take a ``shifted cotangent bundle'' $T^*[-1] \RR\Hol(X,V)$,
and this derived moduli space encodes solutions to
a field theory known as the $\beta\gamma$ system with target~$V$.\footnote{The name is silly: it's due to the original notation people used.}
Every question one asks about the classical physics of the $\beta\gamma$ theory translates into a question about this moduli space.
Other classes of examples are afforded by holomorphic variants of the AKSZ $\sigma$-models (such as the Poisson $\sigma$-model that appears in Kontsevich's construction of deformation quantization); 
the relevant moduli space being the derived mapping space between complex manifolds.

@@

The standard example of a scalar field theory is the free massless exists on any Riemannian manifold,
which we met in Example~\ref{eg: massless scalar}.
Here the field $\phi$ a smooth function on a manifold $M$,
its action functional is $\int_M \phi \Delta \phi$, where $\Delta$ is the Laplace-Beltrami operator, 
and its equation of motion is $\Delta \phi = 0$.
The moduli space of solutions is then the space of harmonic functions on~$M$

When $M$ is two-dimensional and oriented,
we can view $M$ as equipped with a complex structure.
Every complex-valued harmonic function then decomposes as a sum of a holomorphic and anti-holomorphic function. 
That is, if we complexify the scalar field $\phi$, 
we can view it as a pair of fields $\gamma$ and $\Bar{\gamma}$,
where the equation of motion of $\gamma$ is $\dbar \gamma = 0$ and the equation of motion is $\partial \Bar{\gamma} = 0$.
These components are sometimes called the chiral and anti-chiral ``sectors.''

Notice that we have shifted from a second-order differential operator $\Delta$ to first-order differential operators $\dbar$ and~$\partial$.
Reworking a theory to express its equations of motion using first-order operators is sometimes called the first-order formalism.

The holomorphic component can be seen as arising from the following action functional,
\beqn
\int \beta \wedge \dbar \gamma
\eeqn
where $\beta$ is a $(1,0)$-form on $M$ and should be viewed as a Lagrange multiplier for the field $\del \Bar{\gamma}$,
\owen{I don't understand this use of the term Lagrange multiplier. Let's discuss!}
which we now think about as having its own degrees of freedom. 
Note that this action functional also imposes the equation of motion $\dbar \beta = 0$,
so that $\beta$ should be a holomorphic $1$-form.

This piece of the free boson is the simplest holomorphic field theory; it is often called the $\beta\gamma$-\textit{system}.
It plays a role in superstring theory as the free field theory underlying bosonic ghosts corresponding to worldsheet supersymmetry~\cite{Polchinski}.\owen{More precise location?}

@@

\begin{rmk}
We motivated the $\beta\gamma$ system as encoding a piece of the free boson.
There is likewise a complex-conjugate $\Bar{\beta}\Bar{\gamma}$ system as encoding the antichiral sector of the free boson.
To recover the full boson, one must couple the chiral and anti-chiral sectors.
See \cite{Kapustin,GGW} \owen{Where?} for more details. 
\end{rmk}

@@

It is straightforward to modify the example above to allow super-holomorphic vector bundles,
i.e., holomorphic $\ZZ/2$-graded vector bundles.

@@

There is a special class of fermionic theories that is quite interesting.
Let $X$ be a complex $n$-fold where
\begin{itemize}
\item $n = 2k+1$ is odd and
\item there is a line bundle $L \to X$ such that $L^{\otimes 2} \cong K_X$,
i.e., $L$ is a square root of the canonical bundle.
\end{itemize}
A field is a section
\[
\psi \in \Omega^{0,k}(X,\Pi L),
\]
the action functional is
\[ 
\int_X \psi \wedge \dbar \psi,
\]
and the equation of motion is
\[
\dbar \psi = 0.
\]
We suggest to identify solutions that differ by $\dbar$-exact terms.

From the point of view of derived geometry, the cochain complex
\[
\Omega^{0,\bullet}(X, \Pi L)[k]
\]
presents a derived stack that should be viewed as the moduli of solutions for this theory.

\owen{What do you think of this example?}

@@

%In complex dimension $3$ there is a related theory, called Kodaira--Spencer theory, which only consists of the field $\mu$ describing the complex structure in the above example.
Let $X$ be a complex threefold equipped with a Calabi--Yau structure; we denote by $\Omega \in \Omega^{3,hol}(X)$ the holomorphic volume form.
The primary field of Kodaira-Spencer theory is 
\beqn
\mu \in \Omega^{0,1}(X,\T_X),
\eeqn 
a $(0,1)$-form  on $X$ with values in the holomorphic tangent bundle~$\T_X$.
In contrast to examples above, a constraint is imposed on these sections:
we assume that $\mu$ preserve the holomorphic volume form $\Omega$, meaning 
\[
{\rm div}_\Omega(\mu) = 0,
\]
i.e., its divergence vanishes.
In short, the fields are divergence-free vector fields.
The equation of motion is 
\beqn
\dbar \mu + \frac12 [\mu,\mu]_{SN} = 0 ,
\eeqn
where $[-,-]_{SN}$ denotes here the Schouten--Nijenhuis bracket 
(which, on vector fields, is simply the Lie bracket, and is the natural multilinear extension to polyvector fields).
The reader might recognize this equation as the Maurer--Cartan equation controlling deformations of complex structure $\dbar\rightsquigarrow \dbar + \mu$.
Formulating the action functional to obtain this equation of motion is a little involved, 
and we refer to \cite{CLbcov1} for a systematic discussion.

\begin{rmk}
The kinetic term has the form
\[
\int_X \frac12 \mu\, \dbar \del^{-1} \mu
\]
and hence resembles the chiral sector of the scalar, which we just discussed.
\end{rmk}

>>>>

As a more mundane example, let $\RR\Hol(X,Y)$ denote the derived moduli space of holomorphic maps from some complex manifold $X$ to another complex manifold $Y$.
(For $Y = V$ a vector space, a ``derived'' enhancement is familiar as the higher dimensional cohomology $H^*(X, \cO) \otimes V$.)
In derived geometry one can take a ``shifted cotangent bundle'' $T^*[-1] \RR\Hol(X,V)$,
and this derived moduli space encodes solutions to
a field theory known as the $\beta\gamma$ system with target~$V$.\footnote{The name is silly: it is due to the original notation people used.}
Every question one asks about the classical physics of the $\beta\gamma$ theory translates into a question about this moduli space.
Other classes of examples are afforded by holomorphic variants of the AKSZ $\sigma$-models (such as the Poisson $\sigma$-model that appears in Kontsevich's construction of deformation quantization); 
the relevant moduli space being the derived mapping space between complex manifolds.

@@

We met the standard example of a scalar field theory in Example~\ref{eg: massless scalar},
where the field $\phi$ a smooth real-valued function on a manifold $M$
and the equation of motion is $\Delta \phi = 0$,
so that the moduli space of solutions is the space of harmonic functions on~$M$.

A natural analogue is to have the field be a complex-valued function $\gamma$ on a complex manifold $M$ with an equation of motion $\dbar \gamma = 0$,
so that the moduli space of solutions is the space of holomorphic functions on~$M$.

Notice that we are shifting from a second-order differential operator $\Delta$ to first-order differential operators $\dbar$.
Reworking a theory to express its equations of motion using first-order operators is sometimes called the first-order formalism.

We will start with $M$ being a Riemann surface for simplicity.
Consider the action functional
\beqn
\label{eqn: 1d betagamma}
\int \beta \wedge \dbar \gamma
\eeqn
where $\beta$ is a $(1,0)$-form on $M$.\footnote{It can be viewed as a Lagrange multiplier for the field $\del \Bar{\gamma}$, 
which we now think about as having its own degrees of freedom. 
\owen{I don't understand this use of the term Lagrange multiplier. Let's discuss!}}
Note that this action functional also imposes the equation of motion $\dbar \beta = 0$,
so that $\beta$ should be a holomorphic $1$-form.

@@

\begin{rmk}
It is possible to directly relate the usual massless scalar field theory to this $\beta\gamma$-system as follows.

When $M$ is two-dimensional, oriented and Riemannian,
we can view $M$ as equipped with a complex structure.
Every complex-valued harmonic function then decomposes as a sum of a holomorphic and anti-holomorphic function. 
That is, if we complexify the scalar field $\phi$, 
we can view it as a pair of fields $\gamma$ and $\Bar{\gamma}$,
where the equation of motion of $\gamma$ is $\dbar \gamma = 0$ and the equation of motion is $\partial \Bar{\gamma} = 0$.
These components are sometimes called the chiral and anti-chiral ``sectors.''

The $\beta\gamma$-system also plays a role in superstring theory as the free field theory underlying bosonic ghosts corresponding to worldsheet supersymmetry~\cite{Polchinski}.\owen{More precise location?}
There is likewise a complex-conjugate $\Bar{\beta}\Bar{\gamma}$-system encoding the antichiral sector of the free boson.
To recover the full boson, one must couple the chiral and anti-chiral sectors.
See \cite{Kapustin,GGW} \owen{Where?} for more details.~\hfill$\Diamond$
\end{rmk}

@@

It is straightforward to modify the example above to allow holomorphic {\it super} vector bundles,
i.e., holomorphic $\ZZ/2$-graded vector bundles.

@@

There is a special class of fermionic theories that is quite interesting.
Let $X$ be a complex $n$-fold where
\begin{itemize}
\item $n \equiv 1 \mod 4$ (i.e., $n=4k+1$ for some $k$) and
\item there is a line bundle $L \to X$ such that $L^{\otimes 2} \cong K_X$,
i.e., $L$ is a square root of the canonical bundle.
\end{itemize}
A field is a section
\[
\psi \in \Omega^{0,2k}(X,\Pi L),
\]
the action functional is
\[ 
\int_X \psi \wedge \dbar \psi,
\]
and the equation of motion is
\[
\dbar \psi = 0.
\]
We suggest to identify solutions that differ by $\dbar$-exact terms.

From the point of view of derived geometry, the cochain complex
\[
\Omega^{0,\bullet}(X, \Pi L)[2k]
\]
presents a derived super stack that should be viewed as the moduli of solutions for this theory.

\owen{I think I got the dimension condition correct. Can you double-check?}

@@

%In complex dimension $3$ there is a related theory, called Kodaira--Spencer theory, which only consists of the field $\mu$ describing the complex structure in the above example.
Let $X$ be a complex threefold equipped with a Calabi--Yau structure; we denote by $\Omega \in \Omega^{3,hol}(X)$ the holomorphic volume form.
The primary field of Kodaira-Spencer theory is 
\beqn
\mu \in \Omega^{0,1}(X,\T_X),
\eeqn 
a $(0,1)$-form  on $X$ with values in the holomorphic tangent bundle~$\T_X$.
In contrast to examples above, a constraint is imposed on these sections:
we assume that $\mu$ preserve the holomorphic volume form $\Omega$, meaning 
\[
{\rm div}_\Omega(\mu) = 0,
\]
i.e., its divergence vanishes.
In short, the fields are divergence-free vector fields.
The equation of motion is 
\beqn
\dbar \mu + \frac12 [\mu,\mu]_{SN} = 0 ,
\eeqn
where $[-,-]_{SN}$ denotes here the Schouten--Nijenhuis bracket.\footnote{On vector fields, $[-,-]_{SN}$ is simply the Lie bracket, and it is the natural multilinear extension to polyvector fields.}
The reader might recognize this equation as the Maurer--Cartan equation controlling deformations of complex structure $\dbar\rightsquigarrow \dbar + \mu$.
Formulating the action functional to obtain this equation of motion is a little involved, 
and we refer to \cite{CLbcov1} for a systematic discussion.

\begin{rmk}
The kinetic term has the form
\[
\int_X \frac12 \mu\, \dbar \del^{-1} \mu
\]
and hence resembles the chiral sector of the scalar, which we just discussed in the preceding example.
\end{rmk}
